\documentclass[MASTER.tex]{subfiles} 
\begin{document} 
	
	
\begin{frame}
\[ \mbox{Part 2 : Other Interesting Python Packages} \]
\end{frame}

%===========================================================%
\begin{frame}
	\frametitle{statsmodels}
	\begin{itemize}
	\item \texttt{statsmodels} provides a large range of cross-sectional models aswell assometime-series models. 
	\item statsmodels
	uses a model descriptive language (provided via the Python package patsy) to formulate the model
	when working with pandas DataFrames.
	\item Models supported include linear regression, generalized linear
	models, limited dependent variable models, ARMA and VAR models.
	\end{itemize}
\end{frame}
%===========================================================%
\begin{frame}
	\frametitle{Bokeh Data Visualization}
	\begin{itemize}
		\item interactive graphics for the web
		\item designed for large data sets
		\item Designed for streaming data
		\item Native interface in pythin
		\item Fast javascript components
		\item DARPA funded
		\item v.01 relase imminent
	\end{itemize}
	
\end{frame}
%===========================================================%
\begin{frame}
	\frametitle{scikit.learn}
	\begin{itemize}
	\item scikit-learn is an open source machine learning library for the Python programming language. 
	\item It features various classification, regression and clustering algorithms including support vector machines, logistic regression, naive Bayes, random forests, gradient boosting, k-means and DBSCAN, and is designed to interoperate with the Python numerical and scientific libraries NumPy and SciPy.
	\end{itemize}
	.
\end{frame}
%===========================================================%
\begin{frame}
	\frametitle{pytz and babel}
ptyz and babel provide extended support for time zones and formatting information.
\end{frame}
%===========================================================%
\begin{frame}
	\frametitle{rpy2 }
rpy2 provides an interface for calling R 3.0.x in Python, as well as facilities for easily moving data between
the two platforms.
\end{frame}
%===========================================================%
\begin{frame}
	\frametitle{ PyTables and h5py }
PyTables and h5py both provide access to HDF5 files, a flexible data storage format optimized for numeric
data.
\end{frame}
\end{document}